\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-03 gwnnDraft02.tex
refs

The form of the GW energy density at the time of nucleation is given by~\citep{Breitbach:2018ddu},
\begin{equation}\label{eqn::GWengdensity}
h^2\Omega_{\textrm{GW}}^* = 7.7\times 10^{-2} \bigg( \frac{\kappa_{\phi} \alpha}{1 + \alpha} \bigg)^2 \bigg( \frac{H}{\beta} \bigg)^2  S(f)
\end{equation}
where quantities such as $\Omega_{\textrm{GW}}^*$ that are calculated at the time of nucleation are denoted with an asterisk. These must then be evolved to relate to their values at the time of observation. Where  $v \sim 1$ for runaway bubbles. $S(f)$ is the spectral shape function for the signal and a parametric from has been found through numerical simulations~\cite{Huber_2008} of bubble wall collisions.
\begin{equation}\label{eqn::spectralshape}
S(f) = \frac{3.8 \,(f/f_p)^{2.8}}{1 + 2.8\,(f/f_p)^{3.8}}
\end{equation}
The peak frequency, $f_{p}$ is a function of the temperature of the SM at the time of nucleation.  The various hidden sectors can phase transition at different scales and therefore temperatures, causing a shift in the GW spectrum's peak frequency given by~\cite{Huber_2008},
%
\begin{equation}\label{eqn:peakFrequency}
f_{p} = 3.8 \times 10^{-8} \, \textrm{Hz}\, \bigg( \frac{\beta}{H}\bigg)\bigg(\frac{T_{\gamma}}{100 \, \textrm{GeV}}\bigg)\bigg(\frac{g_{*}}{100}\bigg)^{\frac{1}{6}},
\end{equation}
%
where $g_*$  is calculated using Eq.~(\ref{eqn:RelaDOF}) , although, due to the lack of substantial reheating into the hidden sectors, the SM contribution is dominant.

@@

The GW results of these cases are presented in Fig.~\ref{fig:Haa}. In all cases, the summed GW signal is detectable by one or more proposed interferometers. When changing the assumptions on $\beta/H$, the scenarios in Fig.~\ref{fig:Haa} are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$. As $\beta/H$ increases (decreases) the peak frequency moves to higher (lower) frequencies, dictated by Eq.~(\ref{eqn:peakFrequency}), where as the amplitude decreases (increases) shown in Eq.~(\ref{eqn::GWengdensity}).

@@

A \textit{stochastic} gravitational wave background could be detectable if the the signal-to-noise ratio (SNR) is above some threshold value, $\rho > \rho_{th}$, dictated by the capabilities of future interferometers and pulsar timing arrarys (PTA). These interferometers/PTAs quote their experimental sensistivies in terms of spectral noise curves, $S_{\textrm{eff}}(f)$, which can be translated into units of energy density through $h^2 \Omega_{\textrm{eff}}(f) = \frac{2\pi^2}{3H^2}f^3 S_{\textrm{eff}}(f)$. If the experiment uses a single (multiple) detector, the auto-correlated (cross-correlated) SNR is used in comparing to the threshold value $\rho_{th}$. The auto-correlated and cross-correlated SNR are explictely given as~\cite{PhysRevD.59.102001},

@@

\rho^2 =  \mathcal{T} \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\;\;\; \textrm{(auto-correlated)}

@@

\rho^2 = 2 \mathcal{T} \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\: \textrm{(cross-correlated)}

@@

Where $\mathcal{T}$ is the exposure time of the experiment. The integration covers the entire broadband range of frequencies $(f_{\textrm{min}}, f_{\textrm{max}})$. LISA~\citep{Audley:2017drz} and B-DEICIGO~\citep{10.1093/ptep/pty078} are proposed to be single detector interferometers, where as BBO~\citep{PhysRevD.72.083005} and DEICIGO~\citep{Sato_2017} would be built from an array of multiple interferometers.

@@

h^2 \Omega_{\textrm{GW}}(f) = h^2 \Omega_{\gamma}  \bigg(\frac{f}{f_{\textrm{ref}}}\bigg)^{\gamma}

@@

where $\gamma$ is the spectral index of the power law, and $f_{\textrm{ref}}$ is an arbitrary reference frequency which has no effect on the PLI sensetivities. 
The method of calculating the PLI curves involves plotting $h^2 \Omega_{\textrm{GW}}(f) $, using Eq.~(\ref{eqn:RelaDOF}), for various spectral indices $\gamma \in (-8,-7,-6, ... , 6, 7, 8)$ and for some fixed threshold value of $\rho_{th}$. Each curve will lay tangent to the PLI curve, more formally,

@@

h^2 \Omega_{\textrm{PLI}} = \max\limits_{\gamma}\bigg[ h^2 \Omega_{\gamma}\bigg(\frac{f}{f_{\textrm{ref}}}\bigg)^{\gamma}  \bigg].

@@

The spectral noise curves used to create the PLI curves shown in Figs.~\ref{fig::Nnatural} and \ref{fig:Haa} were taken from~\citep{Robson_2019,PhysRevD.83.084036,doi:10.1142/S0218271813410137,10.1093/ptep/pty078, Breitbach:2018ddu} for the interferometers, and~\citep{Breitbach:2018ddu, Janssen:2014dka} for the SKA pulsar timing array. We have assumed an observation time of $\mathcal{T} = 4 $ years for the interferometers and $\mathcal{T}= 5, 10, 20 $  years for the various stages of SKA. In the case of the PTA experiments, the sensitivity curves are dependent on how frequently the pulsar's timing residuals, $\delta t$, are measured. When using Eq.~(\ref{eqn::SNR}) to construct the PSI curves for SKA, the upper integration bound is inversely proportional to pulsar's timing residual, $f_{min} = 1/\delta t$. In this work, it is assumed that $\delta t = 14$ days, but this may underestimate the capabilities of SKA as well as the cadences of the pulsar populations. If the timing residuals are lowered the maximum frequency reach of SKA increases, and the corresponding PSI curves in Figs.~(\ref{fig::Nnatural}) $\&$ ~(\ref{fig:Haa}) are shifted to the right, possibly giving the PTAs sensitivity to some of the scenarios considered here.

@@

Here, we examined several scenarios (including $N$naturalness) that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections demonstrate that although vanilla $N$naturalness is not projected to be detectable in the near future, more generalized scenarios with multiple hidden sector SFOPTs are in an observable region and will begin to be probed by next generation space experiments. Both cases feature important parts of their GW signals in the void between frequencies detectable by pulsar timing arrays and space-based interferometers --- providing theoretical impetus for new experiments capable of probing this region of frequency space.

@@

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. If a model predicts of the presence of 2, 5, or more additional sectors, or features a single extra sector with multiple PTs, deviations from a standard power law can occur. The multiple transitions that occur in the models outlined here create signals that follow this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors. ultimately providing another means to test and constrain theories.

>>>>

The form of the GW energy density at the time of nucleation is given by~\citep{Breitbach:2018ddu},
\begin{equation}
h^2\Omega_{\textrm{GW}}^* = 7.7\times 10^{-2} \bigg( \frac{\kappa_{\phi} \alpha}{1 + \alpha} \bigg)^2 \bigg( \frac{H}{\beta} \bigg)^2  S(f)
\end{equation}
where quantities such as $\Omega_{\textrm{GW}}^*$ that are calculated at the time of nucleation are denoted with an asterisk. These must then be evolved to relate to their values at the time of observation. Where we have assumed \draftnote{isn't this the definition of runaway?} an optimistic bubble wall velocity of $v \sim c$ for runaway bubbles. $S(f)$ is the spectral shape function for the signal and a parametric from has been found through numerical simulations~\cite{Huber_2008} of bubble wall collisions.
\begin{equation}\label{eqn::spectralshape}
S(f) = \frac{3.8 \,(f/f_p)^{2.8}}{1 + 2.8\,(f/f_p)^{3.8}}
\end{equation}
The peak frequency, $f_{p}$ is a function of the temperature of the SM at the time of nucleation.  The various hidden sectors can phase transition at different scales and therefore temperatures, causing a shift in the GW spectrum's peak frequency given by~\cite{Huber_2008},
%
\begin{equation}\label{eqn:peakFrequency}
f_{p} = 3.8 \times 10^{-8} \, \textrm{Hz}\, \bigg( \frac{\beta}{H}\bigg)\bigg(\frac{T_{\gamma}}{100 \, \textrm{GeV}}\bigg)\bigg(\frac{g_{*}}{100}\bigg)^{\frac{1}{6}},
\end{equation}
%
where $g_*$  is calculated using Eq.~(\ref{eqn:RelaDOF}) , although, due to the lack of substantial reheating into the hidden sectors, the SM contribution is dominant.

@@

The GW results of these cases are presented in Fig.~\ref{fig:Haa}. In all cases, the summed GW signal is detectable by one or more proposed interferometers. When changing the assumptions on $\beta/H$, the scenarios in Fig.~\ref{fig:Haa} are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$. \draftnote{Give more details about what happens for changing $\beta$}

@@

A unpolarized stocastic gravitational wave background could be detectable if the the signal-to-noise ratio (SNR) is above some threshold value, $\rho > \rho_{th}$, dictated by the capabilities of future interferometers and pulsar timing arrarys (PTA). These interferometers/PTAs quote their experimental sensistivies in terms of spectral noise curves, $S_{\textrm{eff}}(f)$, which can be translated into units of energy density through $h^2 \Omega_{\textrm{eff}}(f) = \frac{2\pi^2}{3H^2}f^3 S_{\textrm{eff}}(f)$. If the experiment uses a single (multiple) detector, the auto-correlated (cross-correlated) SNR is used in comparing to the threshold value $\rho_{th}$. The auto-correlated and cross-correlated SNR are explictely given as~\cite{PhysRevD.59.102001},

@@

\rho^2 =  T \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\;\;\; \textrm{(auto-correlated)}

@@

\rho^2 = 2 T \int_{f_{\textrm{min}}}^{f_{\textrm{max}}}\textrm{d}f \bigg( \frac{h^2 \Omega_{\textrm{GW}}(f)}{h^2 \Omega_{\textrm{eff}}(f)} \bigg)^2 \;\;\;\: \textrm{(cross-correlated)}

@@

Where $T$ is the exposure time of the experiment. The integration covers the entire broadband range of frequencies $(f_{\textrm{min}}, f_{\textrm{max}})$. LISA~\citep{Audley:2017drz} and B-DEICIGO~\citep{10.1093/ptep/pty078} are proposed to be single detector interferometers, where as BBO~\citep{PhysRevD.72.083005} and DEICIGO~\citep{Sato_2017} would be built from an array of multiple interferometers.

@@

h^2 \Omega_{\textrm{GW}}(f) = h^2 \Omega_{\beta}  \bigg(\frac{f}{f_{\textrm{ref}}}\bigg)^{\beta}

@@

where $\beta$ is the spectral index of the power law, and $f_{\textrm{ref}}$ is an arbitrary reference frequency which has no effect on the PLI sensetivities. 
The method of calculating the PLI curves involves plotting $h^2 \Omega_{\textrm{GW}}(f) $, using Eq.~(\ref{eqn:RelaDOF}), for various spectral indices $\beta \in (-8,-7,-6, ... , 6, 7, 8)$ and for some fixed threshold value of $\rho_{th}$. Each curve will lay tangent to the PLI curve, more formally,

@@

h^2 \Omega_{\textrm{PLI}} = \max\limits_{\beta}\bigg[ h^2 \Omega_{\beta}\bigg(\frac{f}{f_{\textrm{ref}}}\bigg)^{\beta}  \bigg].

@@

The spectral noise curves used to create the PLI curves shown in Figs.~\ref{fig::Nnatural} and \ref{fig:Haa} were taken from~\citep{Robson_2019,PhysRevD.83.084036,doi:10.1142/S0218271813410137,10.1093/ptep/pty078, Breitbach:2018ddu} for the interferometers, and~\citep{Breitbach:2018ddu, Janssen:2014dka} for the SKA pulsar timing array. We have assumed an observation time of $T = 4 $ years for the interferometers and $T= 5, 10, 20 $  years for the various stages of SKA. In the case of the PTA experiments, the sensitivity curves are dependent on how frequently the pulsar's timing residuals, $\delta t$, are measured. When using Eq.~(\ref{eqn::SNR}) to construct the PSI curves for SKA, the upper integration bound is inversely proportional to pulsar's timing residual, $f_{min} = 1/\delta t$. In this work, it is assumed that $\delta t = 14$ days, but this may underestimate the capabilities of SKA as well as the cadences of the pulsar populations. If the timing residuals are lowered the maximum frequency reach of SKA increases, and the corresponding PSI curves in Figs.~(\ref{fig::Nnatural}) $\&$~(\ref{fig:Haa}) are shifted to the right, possibly giving the PTAs sensitivity to some of the scenarios considered here.

@@

Here, we examined scenarios including $N$naturalness that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections demonstrate that although vanilla $N$naturalness is not projected to be detectable in the near future, more generalized scenarios with multiple hidden sector SFOPTs are in an observable region and will begin to be probed by next generation space experiments. Both cases feature important parts of their GW signals in the void between frequencies detectable by pulsar timing arrays and space-based interferometers --- providing theoretical impetus for new experiments capable of probing this region of frequency space.

@@

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. If a model predicts of the presence of 2, 5, or more additional sectors, or features a single extra sector with multiple PTs, deviations from a standard power law can occur. The multiple transitions that occur in the models outlined here create signals that follow this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors, ultimately providing another means to test and constrain theories.
